High-Dimensional Regularized Discriminant Analysis

### 2013-05-13 reports/paper-hdrda.Rnw
Fixed an alignment issue.

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
the decision rule in \eqref{eq:hdrda} is equivalent to
	\begin{align}
		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
	\end{align}
\end{thm}
\begin{proof}
From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(
\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U^{T}$ and $|\tilde{\bm \Sigma}_k| =
\gamma^{p-q}| \bm W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $\bm x \in
\mathbb{R}_{p \times 1}$, we have that
	\begin{align*}
	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k | + (p - q) \log \gamma.
	\end{align*}
Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
$\bm U_2^{T} (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
\end{proof}

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$ and required a larger
subset of the features to perform competitively in our study.
\begin{figure}
<<error_rates, fig.width=8, fig.height=8, out.width='.65\\linewidth'>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(color = Classifier), size = 1, linetype = 2)
p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 4)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
p <- p + theme(axis.text = element_text(size = 10))
p <- p + theme(axis.title.x = element_text(size = 17))
p <- p + theme(axis.title.y = element_text(size = 17))
p <- p + theme(strip.text.x = element_text(size = 16))
p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
p <- p + theme(legend.text = element_text(size = 15))
p <- p + theme(legend.key.size = unit(1, "cm"))
p
@ 
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. Approximate standard errors are no
  greater than \Sexpr{max_se}.}
\label{fig:error-rates}
\end{figure}
Next, we examine the optimal estimates of the tuning parameters for the
\emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
optimal estimates of $\lambda$ and $\gamma$ in \ref{suppB}. For
each of the four data sets, the distribution of the optimal values of $\lambda$
selected through cross-validation often had a notable bathtub shape, resembling
a Beta distribution with both shape parameters less than unity. Specifically,
the optimal values of $\lambda$ were typically near 0 or near 1, corresponding
to quadratic and linear discriminant analysis, respectively. For instance, with
smaller values of $d$ applied to the Shipp data set, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in the \emph{HDRDA} classifier preferring the more
parsimonious assumption of equal covariance matrices. In this instance we can
see the flexibility of our proposed classifier to employ the more realistic
assumption that the class covariance matrices are unequal, but if optimal, we
can employ the more parsimonious assumption of equal covariance matrices.

>>>>

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
the decision rule in \eqref{eq:hdrda} is equivalent to
	\begin{align}
		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
	\end{align}
\end{thm}
\begin{proof}
From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(
\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U^{T}$ and $|\tilde{\bm \Sigma}_k| =
\gamma^{p-q}| \bm W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $\bm x \in
\mathbb{R}_{p \times 1}$, we have that
	\begin{align*}
	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k |\\
  &+ (p - q) \log \gamma.
	\end{align*}
Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
$\bm U_2^{T} (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
\end{proof}

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$ and required a larger
subset of the features to perform competitively in our study.
\begin{figure}
<<error_rates, fig.width=8, fig.height=8, out.width='.65\\linewidth'>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(color = Classifier), size = 1, linetype = 2)
p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 3)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
p <- p + theme(axis.text = element_text(size = 10))
p <- p + theme(axis.title.x = element_text(size = 17))
p <- p + theme(axis.title.y = element_text(size = 17))
p <- p + theme(strip.text.x = element_text(size = 16))
p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
p <- p + theme(legend.text = element_text(size = 15))
p <- p + theme(legend.key.size = unit(1, "cm"))
p
@ 
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. Approximate standard errors are no
  greater than \Sexpr{max_se}.}
\label{fig:error-rates}
\end{figure}
Next, we examine the optimal estimates of the tuning parameters for the
\emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
optimal estimates of $\lambda$ and $\gamma$ in \ref{suppB}. For
each of the four data sets, the distribution of the optimal values of $\lambda$
selected through cross-validation often had a notable bathtub shape, resembling
a Beta distribution with both shape parameters less than unity. Specifically,
the optimal values of $\lambda$ were typically near 0 or near 1, corresponding
to quadratic and linear discriminant analysis, respectively. For instance, with
smaller values of $d$ applied to the Shipp data set, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in the \emph{HDRDA} classifier preferring the more
parsimonious assumption of equal covariance matrices. In this instance we can
see the flexibility of our proposed classifier to employ the more realistic
assumption that the class covariance matrices are unequal, but if optimal, we
can employ the more parsimonious assumption of equal covariance matrices.
